### 2015-12-23 paper.tex
Remove McMasters figure

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach any log files that you could have,
% and specify the details of your LaTeX setup (which operating system and LaTeX
% version / tools you are using).

%=================================================================

% LaTeX Class File and Rendering Mode (choose one)
% You will need to save the "mdpi.cls" and "mdpi.bst" files into the same folder
% as this template file.

%=================================================================

\documentclass[energies,article,accept,moreauthors,pdftex,12pt,a4paper]{mdpi}

%=================================================================
\setcounter{page}{1}
\lastpage{x}
\doinum{10.3390/------}
\pubvolume{xx}
\pubyear{2015}
\history{Received: xx / Accepted: xx / Published: xx}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to
% arXiv.org
%\pdfoutput=1

%=================================================================

% Add packages and commands to include here
% The amsmath, amsthm, amssymb, hyperref, caption, float and color packages are
% loaded by the MDPI class.
\usepackage{graphicx}
%\usepackage{subfigure,psfig}
\usepackage[draft]{todonotes}
\usepackage{lineno}
\usepackage[utf8]{inputenc}

\def \p{\partial}
\def \d{\mathrm{d}}
\def \D{\mathrm{D}}

%=================================================================
%% Please use the following mathematics environments:
%\theoremstyle{mdpi}
%\newcounter{thm}
%\setcounter{thm}{0}
%\newcounter{ex}
%\setcounter{ex}{0}
%\newcounter{re}
%\setcounter{re}{0}
%\newtheorem{Theorem}[thm]{Theorem}
%\newtheorem{Lemma}[thm]{Lemma}
%\newtheorem{Characterization}[thm]{Characterization}
%\newtheorem{Proposition}[thm]{Proposition}
%\newtheorem{Property}[thm]{Property}
%\newtheorem{Problem}[thm]{Problem}
%\newtheorem{Example}[ex]{Example}
%\newtheorem{Remark}[re]{Remark}
%\newtheorem{Corollary}[thm]{Corollary}
%\newtheorem{Definition}[thm]{Definition}
%% For proofs, please use the proof environment (the amsthm package is loaded by
% the MDPI class).

%=================================================================

% Full title of the paper (Capitalized)
\Title{Effects of Reynolds Number on the Energy Conversion and Near-Wake
    Dynamics of a High Solidity Vertical-Axis Cross-Flow Turbine}

% Authors (Add full first names)
\Author{Peter Bachant $^{1,}$* and Martin Wosnik $^{1}$*}

% Affiliations / Addresses (Add [1] after \address if there is only one
% affiliation.)
\address{%
$^{1}$ Center for Ocean Renewable Energy, University of New Hampshire, 24
Colovos Rd., Durham, NH, USA}

% Contact information of the corresponding author (Add [2] after \corres if
% there are more than one corresponding author.)
\corres[2]{pxL3@unh.edu, martin.wosnik@unh.edu}

\linenumbers

% Abstract (Less than 200 words)
\abstract{Experiments were performed with a large laboratory-scale high solidity
    cross-flow turbine to investigate Reynolds number effects on performance and
    wake characteristics, to establish scale thresholds for physical and numerical
    modeling of individual devices and arrays. It was demonstrated that the
    performance of the cross-flow turbine becomes essentially $Re$-independent at a
    Reynolds number based on turbine diameter of $Re_D \approx 10^6$, or an
    approximate average Reynolds number based on blade chord of $Re_c \approx 2
    \times 10^5$. A simple model that calculates peak torque coefficient from static
    foil data and cross-flow turbine kinematics was shown to be a reasonable
    estimator for predicting Reynolds number dependence of an actual cross-flow
    turbine operating under dynamic conditions. Mean velocity and turbulence
    measurements in the near-wake showed subtle differences over the range of $Re$
    investigated. However, when transport terms for the streamwise momentum and mean
    kinetic energy were calculated, a similar $Re$ threshold was revealed. These
    results imply that physical model studies of cross-flow turbines should achieve
    $Re_D \sim 10^6$ to properly approximate both the performance and wake dynamics
    of full scale devices and arrays.}

% Keywords: add 3 to 10 keywords
\keyword{Reynolds number; cross-flow turbine; turbine performance; marine
hydrokinetic energy; wind energy; VAWT; scale model}

% The fields PACS, MSC, and JEL may be left empty or commented out if not
% applicable
%\PACS{}
%\MSC{}
%\JEL{}

\begin{document}

\section{Introduction}

Scaled physical models are often used in science and engineering to approximate
real-world systems. The principle of dynamic similarity allows for geometrically
scaled systems to be dynamically identical if certain nondimensional physical
parameters are matched. In the case of fluid systems, the most important
nondimensional parameter is often the Reynolds number, $Re$, which quantifies
the importance of inertial forces over viscous forces on the flow physics
\cite{Acheson1990}: $Re \equiv Ul/\nu$, where $U$ and $l$ are characteristic
velocity and length scales, respectively, and $\nu$ is the fluid kinematic
viscosity. The other common dynamical scale---for systems with a free
surface---is the Froude number $Fr \equiv U/\sqrt{gl}$, where $g$ is the
gravitational acceleration. Matching both Froude and Reynolds number is
difficult for the same fluid, which is illustrated by the relation $Re = l^{3/2}
g^{1/2} Fr / \nu$, since $Re$ scales linearly with $l$, or the geometric scale.
In general the geometric scale of a physical model is not necessarily the same
as its dynamical scale. As such, hereafter we will use the word ``scale'' to
refer to this dynamical scale or Reynolds number and assume Froude number is
small enough to be negligible.

With regards to wind and marine hydrokinetic (MHK) turbines, scaled physical
models are used to validate predictive numerical models, predict full-scale
performance of individual turbines, and design or investigate arrays of devices.
Scaled models have the benefit of being significantly less expensive, however, a
key question is whether or not an acceptable scale mismatch exists. Similarly,
it is questionable whether numerical models validated with physical model data
obtained orders of magnitude away from prototype scale should be considered
validated at all. An example of the errors that can result from the
extrapolation of small scale experiments can be found in \cite{Baker1991}.

The performance and wake characteristics of cross-flow (vertical-axis) turbines
depend on turbine solidity, blade profile (lift/drag, dynamic stall at reduced
frequency of turbine rotation, symmetry, thickness, camber), blade pitch, number
of blades, strut drag and operational parameters such as tip speed ratio---{\em
    and} on Reynolds number \cite{Para2002}. Note that an average blade chord
Reynolds number, $Re_{c,\mathrm{avg}} \approx \lambda U_\infty c/ \nu$, can be
expressed in terms of tip speed ratio $\lambda \equiv \omega R/ U_\infty$, where
$U_\infty$ is the free stream velocity, $c$ is the blade chord length, $\nu$ is
the fluid kinematic viscosity, $\omega$ is the rotor's angular velocity, and $R$
is the rotor radius. The value of $\lambda$ at which a turbine reaches peak
performance in general decreases with turbine solidity $Nc/(\pi D)$
\cite{Templin1974}, which allows for the use of a simpler Reynolds number based
on turbine diameter $Re_D \equiv U_\infty D/\nu$.

Solidity often directly correlates with the chord-to-radius ratio $c/R$. Rotors
with $c/R \ge 0.1$ are considered high solidity \cite{Fiedler2009}, for which
so-called flow curvature or virtual camber effects become significant
\cite{Migliore1980}. These effects arise from the blade sections' circular
paths, and can complicate the comparison with behavior of an airfoil in a linear
flow.

Blackwell et al. \cite{Blackwell1976} investigated the effects of Reynolds
number on the performance of a 2 m diameter Darrieus vertical-axis cross-flow
wind turbine with NACA 0012 blades. By varying wind tunnel speed, the turbine
was made to operate at approximately constant blade chord Reynolds number $Re_c$
ranging from $1 \times 10^5$ to $3 \times 10^5$. In this regime the turbine
power coefficient $C_P$ was shown to be sensitive to $Re_c$, with sensitivity
diminishing at the higher Reynolds numbers, especially for turbines of lower
solidity ($Nc/R$, where $N$ is the number of blades and $R$ is the turbine's
maximum radius). More recently, Polagye and Cavagnaro \cite{Polagye2013b}
observed significant Reynolds number sensitivity for a high solidity helical
cross-flow turbine, and Bravo et al. \cite{Bravo2007} observed power coefficient
of a straight-bladed VAWT to become Reynolds number independent at $Re_c \approx
4 \times 10^5$. Bachant and Wosnik \cite{Bachant2014} observed Reynolds number
independence of power coefficient at optimal tip speed ratio for a high solidity
cross-flow turbine at $Re_D \approx 10^6$ or $Re_{c,\mathrm{avg}} \approx 2
\times 10^5$.

The wake of a 2-D cross-flow turbine in the dynamic stall regime has been
studied via laser Doppler velocimetry by Brochier et al. \cite{Brochier1986} and
in 3-D with particle image velocimetry (PIV) by Fujisawa and Shibuya
\cite{Fujisawa2001}. However, both of these studies were performed at very low
Reynolds numbers---$Re_D = 10^4$ and $10^3$, respectively. Tescione et al.
\cite{Tescione2014} studied the wake of a vertical-axis wind turbine at optimal
tip speed ratio in a wind tunnel using stereo PIV at an approximate blade chord
Reynolds number $Re_c = 1.7 \times 10^5$; very close to the $Re$-independence
criteria reported in \cite{Bachant2014}, though it was not confirmed if this was
indeed a $Re$-independent state. The two lower $Re$ experiments tended to focus
on the effects of dynamic stall, whereas the higher $Re$ experiment mostly
concerned the mean velocity and tip vortex effects. Whereas the value of these
studies was in elucidating the complex wake dynamics of cross-flow turbines,
they also motivated the more systematic investigation of scale effects
undertaken here.

Scale effects on axial-flow or horizontal-axis wind turbines are better
understood, and investigators have methods for compensating. Krogstad and
Adaramola \cite{Krogstad2012a} observed Reynolds number independence of the
performance of a 0.9 m diameter axial-flow turbine rotor at $Re_D \approx 5
\times 10^5$ in wind tunnel tests. Their turbine blades had an S826 profile
along their entire span, a profile chosen for its $Re$-independence. Walker et
al. \cite{Walker2014} similarly chose a NACA 63-618 foil for their axial-flow
turbine tests in a towing tank, which were performed at a mid-span Reynolds
number $Re_c = 4 \times 10^5$. McTavish et al. \cite{McTavish2013} showed how
the near-wake expansion for an axial-flow rotor is increased at higher Reynolds
numbers, concluding that physical models should be designed with
$Re$-independence in mind if they are to be run at low $Re$. It is uncertain,
however, if these methods work equally well for cross-flow turbines.

When designing or studying arrays, it is common to use very small
(geometrically) scaled devices \cite{Chamorro2011, Chamorro2011b}. It is
therefore important to realize the limitations of evaluating both the power
output of turbines and the wake recovery when Reynolds number is very far from
full scale. Sometimes models used are not turbines, but wake generating objects,
e.g., porous disks, that are meant to replicate the wakes of real turbines
\cite{Goldenberg1983}. In this case, it is of interest to determine at what
Reynolds number one might be able to realistically study wake flows in an array,
and also to evaluate the effectiveness of a wake generator. In other words, a
wake generator may do a fine job simulating a scaled turbine, but how well can
it simulate a full-scale device? Note that for a porous disk, the Reynolds
number of interest is diameter-based, since it is the scaling of the far-wake
dynamics that matters.

Vertical-axis wind turbine array field experiments have revealed that improved
wake recovery allows for closer spacing when compared to conventional axial-flow
propeller-type turbines \cite{Dabiri2011, Kinzel2012}. It was observed
experimentally that a high solidity vertical-axis cross-flow turbine's near-wake
produces a unique vertical mean velocity field---generated by blade tip vortex
shedding---the advection by which is the largest contributor to streamwise
momentum and energy transport or recovery \cite{Bachant2015-JoT}. In this study
we seek to replicate those same momentum and energy balance considerations at
multiple Reynolds numbers, to examine the implications on how scaled, i.e., low
Reynolds number experiments may be used to study flows in turbine arrays.

\subsection{Modes of Reynolds Number Dependence}

We are looking at how Reynolds number scaling affects both the blade loading,
i.e., turbine performance, and the near-wake. Typically, static airfoil data
shows that the static stall angle increases with blade chord Reynolds number
$Re_c$ \cite{Jacobs1937}. A review of Reynolds number effects on airfoil
behavior is presented in \cite{Lissaman1983}. In general, airfoil
performance---often characterized by the profile's lift-to-drag ratio---is
enhanced as the boundary layer on the foil transitions to turbulence closer to
the leading edge, which enables it to advance further downstream against the
adverse pressure gradient on the suction side, delaying separation to higher
angles of attack. For smooth airfoils, this transition can cause a dramatic
increase in foil performance at a blade chord Reynolds number on the order of
$10^5$ \cite{McMasters1980}. Note that there is a distinct lack of highly
reliable data for airfoils in this transitional regime and below. An evaluation
of the various databases relevant to cross-flow turbines is presented in
\cite{Bedon2014}.

Static foil performance does not tell the whole story for a cross-flow turbine.
The azimuthal, and therefore temporal variation of $\alpha$ in a cross-flow
turbine implies dynamic loading, encountering dynamic stall for tip speed ratios
near and below those of maximum rotor torque \cite{Para2002}. Bousman
\cite{Bousman2000-evaluation} states that dynamic stall is relatively
insensitive to Reynolds number for $Re=1.0 \times 10^5$--$2.5 \times
10^5$---judging from measurements on a pitching VR-7 foil in a wind and water
tunnel---since the loading is dominated by vortex shedding. However, Singleton
and Yeager Jr. \cite{Singleton2000} state that the effect of Reynolds number on
dynamic stall remains an unsolved question.

Despite lower lift on the blades at lower $Re$, we expect stronger tip vortex
shedding \cite{Yoon2005}. As mentioned previously, the dynamic stall vortex
shedding is not expected to be highly sensitive to Reynolds number, though a
larger separation bubble at lower $Re$ may induce higher levels of turbulence as
shed vortices become unstable and break down.

Chamorro et al. \cite{Chamorro2012} showed that turbulence statistics in an AFT
wake became $Re$-independent at $Re_D \approx 1 \times 10^5$, and that mean
velocity profiles became $Re$-independent slightly earlier at $Re_D \approx 5
\times 10^4$. Note that in this study the tip chord Reynolds number was not
reported, but is estimated to be $Re_c \approx 4 \times 10^4$ at $Re_D=1 \times
10^5$. It is therefore our objective to observe similar scaling relationships
for a cross-flow turbine near-wake.

\section{Experimental Setup}

Experiments were performed in a turbine test bed specifically designed for
cross-flow turbines. The test bed was integrated as part of the UNH tow tank,
which is a 36 m long facility with a 3.66 m wide and 2.44 m deep cross-section.
The turbine model used in this study was the UNH Reference Vertical Axis Turbine
(RVAT), which was designed to be a generic case for numerical model testing,
similar to the Sandia National Labs/US Department of Energy RM2 River Turbine
\cite{Neary2014}, but with higher solidity, or blade chord-to-radius ratio.

The UNH-RVAT turbine has three blades made from NACA 0020 profiles with 0.14 m
chord. The blades are mounted at mid-chord w.r.t. the turbine axis, and the
turbine has a height (blade span) of 1 m and a diameter of 1 m, c.f.
Figure~\ref{fig:turbine}. The blockage ratio produced by the rotor's frontal
area is 11\%, which we decided not to correct for, as reliable methods are
not yet agreed upon for CFTs \cite{Cavagnaro2014}. As such, blockage should be
taken into account when using these data for model validation, i.e., the
experimental domain should be mimicked. The rotor has a relatively high solidity
$Nc/(\pi D) = 0.13$ and large chord-to-radius ratio $c/R = 0.28$. A CAD model of
the turbine is available from \cite{Bachant2014-RVAT-CAD}.

\begin{figure}[ht!]
\centering

\includegraphics[width=0.4\textwidth]{figures/turbine}

\caption{UNH-RVAT turbine model. Turbine blades and struts made from NACA 0020
    profiles with 0.140 m chord. Note that the upper and lower mounting flanges have
    been excluded as these were included in the tare drag measurements.}

\label{fig:turbine}
\end{figure}

The turbine was mounted in a frame constructed from NACA 0020 sections, shown in
Figure~\ref{fig:exp-setup}. The turbine shaft ran up through the water surface,
coupled to a Kollmorgen AKM permanent magnet servo motor with a 20:1 gearbox,
providing precise control over shaft RPM. This servo was controlled by the tow
tank's main motion controller for high synchronization with the carriage motion,
thereby giving precise measurement and control of tip speed ratio. An Interface
T8 200 Nm capacity rotary torque transducer was installed inline between the
servo and the turbine, and the servo was also mounted on a slewing ring bearing,
which allowed a redundant measurement of torque via an arm and load cell used to
counteract the turbine moment. The frame was mounted to the carriage via linear
guides, such that the total streamwise drag force was transferred to a pair of
Sentran ZB3 500 lbf capacity S-beam load cells, providing the rotor drag
measurements, after a separately measured tare drag was subtracted in post
processing. Similarly, a tare torque was measured by rotating the turbine shaft
in air. Turbine angular and tow carriage linear position were measured using
quadrature encoder signals, with $10^5$ counts-per-rev for the turbine and 10
$\mathrm{\mu m}$ resolution for the carriage position. These signals, along with
the torque and drag signals, were sampled at 2 kHz.

\begin{figure}[ht!]
    \centering

    \includegraphics[width=0.8\textwidth]{figures/exp-setup-photo}

    \includegraphics[width=0.8\textwidth]{figures/exp_setup_drawing}

    \caption{Experimental setup: turbine test bed installed in UNH tow tank.}

    \label{fig:exp-setup}
\end{figure}

Wake velocity was measured using a Nortek Vectrino+ acoustic Doppler velocimeter
(ADV), which has a 6 mm diameter sampling volume, and sampled at 200 Hz. The
probe was mounted on an automated positioning system, also controlled by the tow
tank's main motion controller. The ADV and DAQ systems' sampling times were
synchronized by triggering the start of data acquisition via a pulse sent from
the motion controller. Additional details of the turbine and experimental setup
are described in \cite{Bachant2015-JoT}.

\subsection{Test Plan}

Approximately 1500 tows were conducted for the study reported here, each tow was
used for a single data point on either a performance curve or wake map. Each
performance curve consisted of 31 tows, where during each tow the mean turbine
tip speed ratio was held constant, ranging from 0.1--3.1 in 0.1 increments. Full
performance curve data were acquired for tow speeds from 0.4 to 1.2 m/s in 0.2
m/s increments, for which turbine diameter and approximate blade chord Reynolds
number are presented in Table~\ref{tab:Re}. Performance was also measured for
$\lambda=1.9$ at tow speeds $[0.3, 0.5, 0.7, 0.9, 1.1, 1.3]$ m/s for two tows
each.

\begin{table}
\centering
\begin{tabular}{ccc}
Tow speed (m/s) & $Re_D$ & $Re_{c,\mathrm{ave}}$ ($\lambda = 1.9$) \\
\hline
0.3 & $0.3 \times 10^6$ & $0.8 \times 10^5$ \\
0.4 & $0.4 \times 10^6$ & $1.1 \times 10^5$ \\
0.5 & $0.5 \times 10^6$ & $1.3 \times 10^5$ \\
0.6 & $0.6 \times 10^6$ & $1.6 \times 10^5$ \\
0.7 & $0.7 \times 10^6$ & $1.9 \times 10^5$ \\
0.8 & $0.8 \times 10^6$ & $2.1 \times 10^5$ \\
0.9 & $0.9 \times 10^6$ & $2.4 \times 10^5$ \\
1.0 & $1.0 \times 10^6$ & $2.7 \times 10^5$ \\
1.1 & $1.1 \times 10^6$ & $2.9 \times 10^5$ \\
1.2 & $1.2 \times 10^6$ & $3.2 \times 10^5$ \\
1.3 & $1.3 \times 10^6$ & $3.4 \times 10^5$ \\
\end{tabular}
\caption{Turbine diameter and approximate blade chord Reynolds numbers for the
tow speeds used in the experiment.}
\label{tab:Re}
\end{table}

Each wake map was generated by positioning a Nortek Vectrino+ acoustic Doppler
Velocimeter (ADV), measuring three components of velocity at 200 Hz sampling
frequency, at 270 different locations, varied in the cross-stream and vertical
directions at one turbine diameter downstream ($x/D=1$). These locations have
vertical coordinates from the turbine centerline up to $z/H=0.625$, ranging in
the cross-stream direction $y/R = \pm 3$. These locations are shown in
Figure~\ref{fig:wake-locations}.

\begin{figure}
\centering

\includegraphics[width=0.9\textwidth]{figures/turbine_coordinate_system}

\caption{Wake measurement coordinate system and locations. Dimensions are in
    meters.}

\label{fig:wake-locations}
\end{figure}

\subsection{Data Processing}

From each set of tows, a standard time interval was set, which allowed the
turbine performance and wake to reach a quasi-periodic state. Each run was
analyzed to compute statistics over this interval, truncating the end slightly
to achieve and integer number of blade passages. Turbine shaft angular velocity
and tow carriage speed were calculated using as second order central
differencing scheme on the respective position measurements. Power and drag
coefficients are calculated as instantaneous quantities using the carriage speed
as the free stream velocity.

Wake velocity data were filtered for spurious ``spikes'' by removing datapoints
8 standard deviations or 0.9 m/s from the mean. The experimental data and code
for processing and visualization are available from
\cite{Bachant2015-RVAT-Re-dep-data}.

\subsection{Uncertainty Analysis}

Uncertainty was considered from a combination of systematic and random errors.
The random error was inferred from the sample standard deviation (on a
per-revolution basis) and the systematic from the sensor calibrations or
datasheets. Combining both sources of error, along with their propagation into
quantities derived from multiple measurements, followed the procedures outlined
in Coleman and Steele \cite{ColemanSteele}, described below.

An expanded uncertainty interval with 95\% confidence was computed for
$C_P$, $C_D$, and mean wake velocities
\begin{equation}
    U_{95} = t_{95} u_c,
\end{equation}
where $t_{95}$ is the value from the Student-$t$ distribution for a 95\%
confidence interval and $u_c$ is the combined standard uncertainty. Combined
standard uncertainty for a given quantity $X$ is calculated by
\begin{equation}
    u_X^2 = s_{\bar{X}}^2 + b_X^2,
\end{equation}
where $s_{\bar{X}}$ is the sample standard deviation of the mean per turbine
revolution, and $b_X$ is the systematic uncertainty, computed by
\begin{equation}
    b_{X}^2 = \sum_{i=1}^J \left( \frac{\partial X}{\partial x_i} \right)^2
    b_{x_i}^2,
\end{equation}
where $x_i$ is a primitive quantity used to calculate $X$ (e.g. $T$, $\omega$,
and $U_\infty$ for calculating $C_P$), and $b_{x_i}$ is the primitive quantity's
systematic uncertainty, estimated as half the value listed on the sensor
manufacturer's documentation.

Selecting $t_{95}$ requires an estimate for degrees of freedom $\nu_X$, which
was obtained using the Welch--Satterthwaite formula
\begin{equation}
    \nu_X = \frac{\left(s_X^2 + \sum_{k=1}^M b_k^2 \right)^2} {s_X^4/\nu_{s_X} +
    \sum_{k=1}^M b_k^4/\nu_{b_k}},
\end{equation}
where $\nu_{s_X}$ is the number of degrees of freedom associated with $s_X$ and
$\nu_{b_k}$ is the number of degrees of freedom associated with $b_k$.
$\nu_{s_X}$ is assumed to be $(N-1)$, where $N$ is the number of independent
samples (or turbine revolutions). $\nu_{b_k}$ was estimated as
\begin{equation}
    \nu_{b_k} = \frac{1}{2} \left( \frac{\Delta b_k}{b_k} \right)^{-2},
\end{equation}
where the quantity in parentheses is the relative uncertainty of $b_k$, assumed
to be 0.25.

For the Reynolds number dependence of turbine performance, error bars are
included on the plots. Expanded uncertainty estimates for mean wake velocities
(averaged over all runs) are listed in Table~\ref{tab:vel-unc}.

\begin{table}
    \centering
\begin{tabular}{cccc}
$U_\infty$ (m/s) &                $U$ &                $V$ &                $W$ \\
\hline
           $0.4$ & $1 \times 10^{-2}$ & $7 \times 10^{-3}$ & $6 \times 10^{-3}$ \\
           $0.6$ & $1 \times 10^{-2}$ & $8 \times 10^{-3}$ & $8 \times 10^{-3}$ \\
           $0.8$ & $2 \times 10^{-2}$ & $1 \times 10^{-2}$ & $1 \times 10^{-2}$ \\
           $1.0$ & $2 \times 10^{-2}$ & $1 \times 10^{-2}$ & $1 \times 10^{-2}$ \\
           $1.2$ & $2 \times 10^{-2}$ & $1 \times 10^{-2}$ & $1 \times 10^{-2}$ \\
\end{tabular}
\caption{Average expanded uncertainty estimates (with 95\% confidence) for mean
    velocity measurements at each tow speed.}

\label{tab:vel-unc}
\end{table}

\section{Results and Discussion}

\subsection{Performance}

Full power and rotor drag (a.k.a. thrust) coefficient curves for various
Reynolds numbers are plotted in Figure~\ref{fig:cp-curves} and
Figure~\ref{fig:cd-curves}, respectively. In general, maximum $C_P$ increases
with Reynolds number. The power coefficient curves also show a slight downward
shift in the optimal tip speed ratio (peak performance) with increasing Reynolds
number, from about $\lambda \approx$ 2.0 to 1.9. This is caused by the stall
delay from a more turbulent boundary layer on the blade suction side. There is
essentially no change in the shape of the drag coefficient curves---merely a
slight upward shift in $C_D$ with increasing $Re$.

\begin{figure}[ht]
\centering

\includegraphics[width=0.65\textwidth]{figures/cp_curves}

\caption{Mean power coefficient curves plotted for multiple Reynolds numbers.}

\label{fig:cp-curves}
\end{figure}

\begin{figure}[ht]
\centering

\includegraphics[width=0.65\textwidth]{figures/cd_curves}

\caption{Mean rotor drag coefficient curves plotted for multiple Reynolds
    numbers.}

\label{fig:cd-curves}
\end{figure}

Mean power and drag coefficients at $\lambda=1.9$ are plotted versus Reynolds
number in Figure~\ref{fig:perf-Re-dep}. Note that the large error bars for $C_P$
at low $Re$ are dominated by systematic error estimates for the torque
transducer, since torque values are at the lower end of the measurement range.
However, the uncertainty due to random error or repeatability remains relatively
low. There is a drastic improvement in $C_P$ with increasing Reynolds number at
lower values of Reynolds number. The power coefficient then becomes  essentially
$Re$-independent at $Re_D = 0.8 \times 10^6$, which corresponds to an
approximate average blade chord Reynolds number $Re_{c, \mathrm{ave}} = 2.1
\times 10^5$. This threshold is consistent with the behavior of the blade
boundary layer transitioning from laminar to turbulent, thereby promoting either
the suppression or reattachment of the laminar separation bubble
\cite{Lissaman1983}.

The behavior of the mean rotor drag coefficient $C_D$ is similar, though the
changes are less dramatic. This is likely due to cross-flow turbine geometry,
where increases in blade drag at lower $Re$ somewhat offsets the reduction in
lift, causing the total streamwise force to vary less than the rotor torque. The
tendency for $C_D$ to continue increasing with $Re$ may also include the effects
of an increasing Froude number, which increases free surface deformation and
wave drag during towing without increasing flow through the turbine.

\begin{figure}[ht]
\centering

\includegraphics[width=0.9\textwidth]{figures/perf_re_dep}

\caption{UNH-RVAT measured mean power (left) and drag (right) coefficients at
    $\lambda=1.9$ plotted versus Reynolds number. Error bars indicate expanded
    uncertainty estimates for 95\% confidence, which for $C_P$ is dominated by
    systematic error estimates from the torque transducer operating at the lower end
    of its measurement range.}

\label{fig:perf-Re-dep}
\end{figure}

\subsubsection{Relation to Static Foil Characteristics}

To help understand---and possibly predict---the $Re$-sensitivity on turbine
performance, a series of static foil coefficient datasets were computed with the
XFOIL viscous panel method code \cite{Drela1989}, a commonly used tool for
airfoil analysis, e.g. \cite{Castelli2011, Walker2014}, implemented as part of
the open source turbine design software QBlade \cite{Marten2013}. Simulations
were run for an angle of attack range of 0--40 degrees, in increments of 0.5
degrees. Solver parameters used were 100 panels, a fixed speed, zero Mach
number, \texttt{NCrit=9} (default $\mathrm{e}^n$ transition criteria parameter
for an average wind tunnel), and no forced boundary layer transition.
Characteristics were computed for the approximate average blade chord Reynolds
numbers encountered in the tow tank experiments.

To investigate the effects of the blades' ``virtual camber'' due to their
circular path \cite{Migliore1980}, the XFOIL calculations were performed for
20\% thick foils with 0\% (NACA 0020), 2\% (NACA 2520), and 4\% (NACA 4520)
camber about their half-chord location. A 4\% camber approximates the maximum
distance between the blade chord line and path for the UNH-RVAT and the 2\%
camber takes into account the reduction in virtual flow curvature from the
non-curved inflow velocity by dividing by the tip speed ratio $\lambda=1.9$.

Results from the XFOIL calculations for the airfoil profiles are shown in
Figure~\ref{fig:foil-Re-dep}, where values of maximum lift coefficient
$C_{l_{\max}}$, minimum drag coefficient $C_{d_{\min}}$, and maximum
lift-to-drag ratio are normalized (for comparison between foils) and plotted
versus $Re_c$. In general, larger camber is associated with decreased foil
performance at lower Reynolds number. The data and processing code for these
calculations is available from \cite{Bachant2015-NACAXX20-XFOIL}.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.9\textwidth]{figures/all_foils_re_dep}
\caption{Foil section characteristics computed by XFOIL for various $Re_c$.}
\label{fig:foil-Re-dep}
\end{figure}

In conjunction with the cross-flow turbine blade kinematics, the foil
coefficients were used to approximate turbine performance by calculating the
peak torque coefficient on the upstream half of the blade path. The turbine
torque coefficient $C_T$ can be related to the blade section chordwise force
coefficient $C_c$ by
\begin{equation}
C_T = \frac{C_c c}{2R} \frac{|U_\mathrm{rel}|^2}{U_\infty^2},
\label{eq:ct}
\end{equation}
where the blade section chordwise force (for zero preset pitch) is given by
\begin{equation}
C_c = C_l \sin \alpha - C_d \cos \alpha.
\label{eq:cc}
\end{equation}

The relative blade velocity $U_\mathrm{rel}$ was calculated by vector addition
of the free stream velocity and the opposite of the blade tangential velocity.
Note that this neglects any induction, or slowing of the free stream by the
turbine forces, which would be present in a momentum/streamtube model. Since the
goal of this approach was not to predict absolute performance but rather gain
insight into relative changes with $Re$, this method was deemed acceptable as it
is extremely simple and fast to compute.

Values for the blade angle of attack, relative velocity, and torque coefficient
are plotted in Figure~\ref{fig:blade-kinematics} for the upstream half of
rotation of the turbine. The effects of static stall are clearly present in the
torque coefficient plot, and by the time the angle of attack has decreased below
that of stall, the relative velocity is so low that there is not much
contribution to the torque coefficient.

\begin{figure}[ht!]
\centering

\includegraphics[width=0.9\textwidth]{figures/foil_kinematics_ct}

\caption{Geometric angle of attack, relative velocity and torque coefficient
    calculated with a NACA 0020 foil operating at $\lambda=1.9$.}

\label{fig:blade-kinematics}
\end{figure}

Results for the normalized peak torque coefficient for each foil are plotted
versus $Re_c$ in Figure~\ref{fig:foils-C_T-Re-dep}. It is interesting that the
convergence of $C_{T_\mathrm{max}}$ with increasing Reynolds number is more
dramatic than any of the common foil performance characteristics plotted in
Figure~\ref{fig:foil-Re-dep}, meaning that the cross-flow turbine's unique
kinematics must be taken into account when attempting to predict the effect of
transitional Reynolds numbers on turbine performance.

From the peak torque coefficient metric plotted in
Figure~\ref{fig:foils-C_T-Re-dep}, the Reynolds number independence is achieved
at lower values, and more dramatically. We see that the trend of the
(non-cambered) NACA 0020 curve matches almost perfectly up to $Re_c \approx 2.1
\times 10^5$, but then continues to increase linearly with a small slope with
$Re$. This is not matched by the experimental data, which at higher $Re$ looks
more like the cambered foil results. Though this method is very approximate, it
nevertheless appears to be a reasonable way to predict the transitional Reynolds
number regime for cross-flow turbine performance using only 2-D static airfoil
characteristics

\begin{figure}[ht!]
\centering

\includegraphics[width=0.65\textwidth]{figures/cft_re_dep_foils}

\caption{Reynolds number dependence of the normalized peak torque coefficient
    calculated from static foil coefficients and blade kinematics, compared to
    experimental data from a cross-flow turbine. Note that the experimental data
    represents the mean torque coefficient, not the maximum.}

\label{fig:foils-C_T-Re-dep}
\end{figure}

\subsection{Wake Characteristics}

The near-wake of this turbine at $x/D=1$, $\lambda=1.9$, and momentum and
kinetic energy balances at a $Re_D = 1.0 \times 10^6$ were discussed in
\cite{Bachant2015-JoT}. Similar data were taken for the experiment here, and
results for the mean velocity field and turbulence kinetic energy calculated
from wake maps of 270 individual measurements (tows) are shown in
Figure~\ref{fig:meancontquiv} and Figure~\ref{fig:kcont}, respectively, looking
upstream towards the turbine. With respect to the mean velocity field, we see
asymmetry and a mean vortex structure created by blade tip vortex shedding. The
effects of the tip vortices are also seen in the turbulence kinetic energy
measurements, along with turbulence generated by the blades undergoing dynamic
stall on the $-y$ side of the turbine.

These same wake maps were measured for $Re_D = 0.4 \times 10^6$, $0.6 \times
10^6$, $0.8 \times 10^6$, and $1.2 \times 10^6$. Qualitatively, these look very
similar, so they have not been plotted here, though profiles at $z/H=0.0$ are
compared in Figure~\ref{fig:profiles} to illustrate the subtlety of the
differences at different $Re$. We will instead compare and contrast the wake
behavior by examining spectra and wake transport terms in the equations that
govern the downstream evolution of mean momentum and kinetic energy.

\begin{figure}[ht!]
    \centering

    \includegraphics[width=0.8\textwidth]{figures/meancontquiv_10}

    \caption{Mean velocity field at $x/D=1$, $\lambda=1.9$, and $Re_D=1.0 \times
        10^6$.}

    \label{fig:meancontquiv}
\end{figure}

\begin{figure}[ht!]
    \centering

    \includegraphics[width=0.8\textwidth]{figures/k_contours_10}

    \caption{Turbulence kinetic energy at $x/D=1$, $\lambda=1.9$, and $Re_D=1.0
        \times 10^6$.}

    \label{fig:kcont}
\end{figure}

\begin{figure}[ht!]
    \centering

    \includegraphics[width=0.85\textwidth]{figures/mean_u_k_profiles}

    \caption{Mean streamwise velocity (left) and turbulence kinetic energy
        (right) profiles at $z/H=0.0$. Legend indicates $Re_D$.}

    \label{fig:profiles}
\end{figure}

\subsubsection{Dominant Timescales and Turbulence Spectra}

Spectra of the cross-stream velocity normalized by the free stream were
computing using a fast Fourier transform-based method, applying a Hanning window
and averaging over four adjacent frequency bands to decrease confidence
intervals. These spectra are plotted in Figure~\ref{fig:wake-spectra} for
regions on either side of the turbine. On the $-y$ side of the turbine there is
broadband turbulence produced by blade stall, and on the $+y$ side there is a
clear peak in the spectra caused by the blade passage. We see that on both sides
there is higher spectral energy at lower Reynolds numbers. On the $+y$ side of
the turbine we notice higher spectral energy at the blade passage frequency's
first harmonic, or $6 f_\mathrm{turbine}$. This is likely due to the blade's
shed vorticity being less stable at higher $Re$.

\begin{figure}[ht!]
\centering

\includegraphics[width=0.85\textwidth]{figures/wake_spectra}

\caption{Cross-stream velocity (normalized by $U_\infty$) spectra at $z/H=0.25$,
    $y/R=-1.0$ ($-0.5$ m in Figure~\ref{fig:wake-locations}) (a) and $y/R=1.5$
    ($+0.75$ m in Figure~\ref{fig:wake-locations}) (b). Dashed vertical lines
    indicate $[1, 3, 6, 9]$ times the turbine rotational frequency. Shaded regions
    indicate 95\% confidence intervals assuming a $\chi^2$ distribution.}

\label{fig:wake-spectra}
\end{figure}

\subsubsection{Transport of Mean Momentum and Kinetic Energy}

The relative importance of various physical processes on mean streamwise
momentum and kinetic energy transport/recovery in the streamwise direction were
assessed by examining the governing equations rearranged to isolate the
streamwise partial derivative ($\partial / \partial x$) for the quantities of
interest \cite{Bachant2015-JoT}, i.e. momentum:

\begin{equation}
	\begin{split}
		\frac{\p U}{\p x}  =  
		\frac{1}{U} \bigg{[}
		& - V\frac{\p U}{\p y}
		- W\frac{\p U}{\p z} \\
		& -\frac{1}{\rho}\frac{\p P}{\p x} \\
		& - \frac{\p}{\p x} \overline{u'u'}
		- \frac{\p}{\p y} \overline{u'v'}
		- \frac{\p}{\p z} \overline{u'w'} \\
		& + \nu\left(\frac{\p^2 U}{\p x^2}
		+ \frac{\p^2 U}{\p y^2}
		+ \frac{\p^2 U}{\p z^2} \right)
		\bigg{]}.
	\end{split}
	\label{eq:momentum}
\end{equation}
and mean kinetic energy:
\begin{equation}
	\begin{split}
		\frac{\p K}{\p x}
		=
		\frac{1}{U}
		\bigg{[}
		& - \underbrace{V \frac{\p K}{\p y}}_{y\text{-adv.}}
		- \underbrace{W \frac{\p K}{\p z}}_{z\text{-adv.}}
		% Pressure work:
		- \frac{1}{\rho}\frac{\p}{\p x_j} P U_i \delta_{ij}
		% Work by viscous forces
		+ \frac{\p}{\p x_j} 2 \nu U_i S_{ij} % Not sure if that's capital S...
		% Turbulent transport of K
		- \underbrace{
		\frac{1}{2}\frac{\p}{\p x_j} \overline{u_i' u_j'} U_i
		}_{\text{Turb. trans.}} \\
		% Production of k
		& + 
		\underbrace{
		\overline{u_i' u_j'} \frac{\p U_i}{\p x_j}
		}_{k\text{-prod.}}
		% Mean dissipation
		- 
		\underbrace{
		2 \nu S_{ij}S_{ij}
		}_{\text{Mean diss.}}
		\bigg{]}.
	\label{eq:K-full}
	\end{split}
\end{equation}

Results from the terms that can be computed from the experimental data (i.e.
excluding $\partial / \partial x$) from the right hand sides of
Equations~\ref{eq:momentum} and \ref{eq:K-full} are plotted in
Figures~\ref{fig:mom-bar-graph} and \ref{fig:K-bar-graph}, respectively.
Derivatives were computed using a second order finite difference scheme, with
central differencing for interior points, and inward facing schemes for the
boundaries. A weighted average for each term was then calculated based on the
grid spacing. We note that similar to \cite{Bachant2015-JoT}, the vertical
advection at this point in the wake is the dominant contributor to positive wake
recovery, caused by the unique vortex pattern created by the blade tips and
wakes.

We see that in general, levels of turbulent transport are slightly lower at
larger Reynolds numbers. The viscous diffusion and dissipation, though still
three orders of magnitude smaller than the other terms, do increase at low
Reynolds numbers, which is consistent with the physical meaning of the Reynolds
number itself. From these results one might expect that the viscous effects will
become significant to the wake dynamics around $Re_D \sim 10^4$.

Transport due to cross-stream advection appears to become more negative at
higher $Re$. This could be a consequence of increasing free surface deformation
(higher $Fr$), which then decreases the effective flow cross-sectional area
downstream of the turbine, forcing more flow to accelerate around the sides of
the turbine. Note that we also see slightly higher drag coefficients at higher
$Re$ (potentially for the same reason), which also helps explain larger negative
values of cross-stream advection.

\begin{figure}[ht!]
    \centering

    \includegraphics[width=0.9\textwidth]{figures/mom_bar_graph}

    \caption{Normalized momentum transport quantities computed as weighted
        averages from Equation~\ref{eq:momentum}.}

    \label{fig:mom-bar-graph}
\end{figure}

\begin{figure}[ht!]
    \centering

    \includegraphics[width=0.9\textwidth]{figures/K_trans_bar_graph}

    \caption{Normalized mean kinetic energy transport quantities computed as
        weighted averages based on Equation~\ref{eq:K-full}. Note that directions of
        the turbulent transport terms refer to the directions of their partial
        derivatives.}

    \label{fig:K-bar-graph}
\end{figure}

The totals for all the wake transport terms calculated in
Figures~\ref{fig:mom-bar-graph} and \ref{fig:K-bar-graph} are plotted in
Figure~\ref{fig:wake-trans-totals}. We see that in general wake transport of
both mean streamwise momentum and kinetic energy is enhanced at lower Reynolds
numbers and levels off consistent with the behavior of the turbine power
coefficient. This is an important consideration if studying sub-scale models of
turbine arrays, where increased levels of wake recovery could motivate different
ideal array configurations when compared with full-scale turbines if the scale
model Reynolds number is too low.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.65\textwidth]{figures/wake_trans_totals}
\caption{Normalized transport totals from Figures~\ref{fig:mom-bar-graph} and
\ref{fig:K-bar-graph} plotted versus Reynolds number.}
\label{fig:wake-trans-totals}
\end{figure}

\section{Conclusions}

In this study it was demonstrated that the performance of a high solidity
($c/R=0.28$) cross-flow turbine becomes essentially $Re$-independent at a
Reynolds number based on turbine diameter of $Re_D \approx 10^6$, or an
approximate Reynolds number based on blade chord of $Re_c \approx 2 \times
10^5$. The power coefficient values on the left side of the $C_P$--$\lambda$
curve are further reduced at low $Re$ due to deeper stalling. The
$Re$-independent threshold corresponds to that at which the foil suction surface
boundary layer becomes turbulent before having to recover static pressure
against the adverse pressure gradient, highlighted in \cite{Lissaman1983,
    McMasters1980, Carmichael1981}.

We propose a method for predicting Reynolds number dependence of a cross-flow
turbine using static airfoil characteristics and turbine blade kinematics:
\begin{enumerate}
    \item Acquire static lift and drag coefficient data for the desired blade
    profile.

    \item For azimuthal angles of 0--180 degrees and a given tip speed ratio,
    calculate the geometric angle of attack and relative velocity magnitude from
    the blade and undisturbed free stream velocity ($\equiv 1$) vectors.

    \item With arrays of geometric angle of attack and relative velocity
    magnitude, calculate the blade chordwise force and then rotor torque
    coefficient from Equations~\ref{eq:cc} and \ref{eq:ct}, respectively.

    \item Extract the maximum value of $C_T$, and repeat this process for static
    foil data at multiple $Re$.
\end{enumerate}

It was shown that the peak torque coefficient computed this way shows similar
Reynolds number sensitivity---a better predictor than conventional
quantifications of airfoil performance, e.g., lift-to-drag ratio. Note that this
method is not restricted to standard symmetrical NACA foils, though those
evaluated here with higher camber were more sensitive at lower $Re$ but had
smaller slopes in the linear regime.

In the near-wake of the turbine we observed lower levels of turbulence with
increasing Reynolds number, along with lower levels of turbulent transport with
respect to mean streamwise momentum and mean kinetic energy recovery in the
streamwise direction. Vertical advection---the largest transport mechanism
measured---showed little $Re$-dependence, whereas the negative effects of
cross-stream advection were enhanced at high $Re$, which may be due to larger
free surface deformation reducing the near-wake's cross-sectional area. Despite
being orders of magnitude smaller than other transport processes, viscous
diffusion increased rapidly with decreasing $Re$, and is expected to become
significant to the wake dynamics around $Re_D \sim 10^4$. Overall, the total
wake transport measured leveled off at essentially the same Reynolds number that
the performance did---$Re_D \approx 0.8 \time 10^6$.

From these results we recommend that physical model tests be performed at $Re_D
\sim 10^6$ to provide reasonable predictions of full-scale performance. This
threshold also applies to the validation of predictive engineering models,
especially for high-fidelity CFD models where the boundary layer is to be
resolved, since transition to turbulence plays an important role in overall
blade loading. Note that in our case blockage, though reasonably low, may be
increasing flow through the turbine compared with a free case, which could mean
$Re$ thresholds for free flows could be slightly higher, though a reliable
blockage correction algorithm has not yet been agreed upon for CFTs
\cite{Cavagnaro2014}.

If using scaled physical models to predict array performance, where turbine
power output is measured, it may also be important to keep all turbines in the
regime where the power coefficient varies linearly to avoid exaggerated power
deficiencies for downstream turbines, despite similarities in wake
characteristics. Results here also suggest that low Reynolds number physical
model studies of turbine arrays may see exaggerated levels of wake recovery,
leading to inadequate or inappropriate spacing or layouts.

\acknowledgements{Acknowledgements}

The authors would like to acknowledge funding through a National Science
Foundation CAREER award (PI Wosnik, NSF-CBET 1150797, program manager Dr.
Gregory L. Rorrer), a grant through the Leslie S. Hubbard Marine Program
Endowment to purchase acoustic flow measurement instrumentation, and a grant for
laboratory infrastructure upgrades through the US Department of Energy.

\conflictofinterests{Conflicts of Interest}

The authors declare no conflict of interest.

\bibliography{library}
\bibliographystyle{mdpi}

\end{document}